---
title: "Stats101A_Final Project_MLR"
author: "Group11"
date: "7/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE}
library(car)
library(gplots)
library(alr3)
library(readr)
library(effects)
stroke <- read.csv("stroke.csv", header=T)
attach(stroke)
```

## Exploratory data analysis

```{r}
WASI<-as.numeric(WASI.Full.Scale...4.Percentile.Rank)
Edu<-as.numeric(Volunteer.Highest.grade.completed.)
BMI<-as.numeric(BMI)
Race<-factor(Race)
High.Cholesterol<-factor(High.Cholesterol)

par(mfrow=c(2,2))
hist(WASI) #y
hist(Edu)
hist(BMI)
table(Race)
table(High.Cholesterol)
par(mfrow=c(2,2))
plot(WASI~Edu,main="The scatterplot between WASI and Education")
plot(WASI~BMI,main="The scatterplot between WASI and BMI")
boxplot(WASI~Race,main="The boxplot between WASI and Race")
boxplot(WASI~High.Cholesterol,main="The boxplot between WASI and High Cholesterol")

summary(Subject.SES.Score) #1Q=40,mdeian=48,3Q=55
SES<- factor(cut(Subject.SES.Score, c(14,42,52,75), c("Low", "Medium", "High")))
table(SES)
boxplot(WASI~SES,main="The boxplot between WASI and SES")
```

## Final model

```{r}
#final model
m1<-lm(WASI~Race+Edu+SES+BMI*High.Cholesterol)
summary(m1)

```

## Correlation matrix and vif

```{r}
cor(stroke[,c(9,19)], use = "complete.obs")
m2<-lm(WASI~Race+Edu+High.Cholesterol+BMI+SES)
vif(m2) #all less than 5

```

## Assumptions

```{r,message=FALSE}
#linearity already checked by scatterplot
#variance constancy
plot(m1,which=1)
ncvTest(m1)

#normality
plot(m1,which=2)

#influential points
influencePlot(m1) #leverage cutoff=4/474=0.00844
stroke_1<- stroke[-c(286,336,296),]
attach(stroke_1)
m3<-lm(WASI~Race+Edu+BMI*High.Cholesterol)
summary(m3) #no difference made by bad leverage points
detach(stroke_1)

#mmp plots
mmps(m1) #fit
```

## AIC
```{r}
#AIC
m01<-lm(WASI~Race)
summary(m01) #R_adj=0.143
AIC(m01) #4402
m02<-lm(WASI~Race+Edu)
summary(m02) #R_adj=0.248
AIC(m02) #4341
m03<-lm(WASI~Race+Edu+BMI)
summary(m03) #R_adj=0.251
AIC(m03) #4339
m04<-lm(WASI~Race+Edu+BMI+High.Cholesterol)
summary(m04) #R_adj=0.265
AIC(m04) #4332
#which one should be chosen?
m1<-lm(WASI~Race+Edu+BMI*High.Cholesterol)
summary(m1) #R_adj=0.272
AIC(m1) #4328
m06<-lm(WASI~Race+Edu+Subject.SES.Score+BMI*High.Cholesterol)
summary(m06) #R_adj=0.277, SES not significant
AIC(m06) #4330

```

## Interaction plot
```{r}
plot(allEffects(m1))
#For people who don't have high cholesterol, as BMI increases, WASI decrease
#For people who have high cholesterol, as BMI increases, WASI also increase
#Why fat people more intelligent?-involve SES score

#Involve with SES(put more than 2 levels categorical variable here)
summary(Subject.SES.Score) #1Q=40,mdeian=48,3Q=55
SES<- factor(cut(Subject.SES.Score, c(14,42,52,75), c("Low", "Medium", "High")))
table(SES)
boxplot(WASI~SES,main="The boxplot between WASI and SES")
m4<-lm(WASI~Race+Edu+BMI*SES)
summary(m4)
plot(allEffects(m4))
#For people who have low SES, as BMI increases, WASI decrease
#Low SES correpond to people who don't have high cholesterol
boxplot(Subject.SES.Score~High.Cholesterol)
tapply(Subject.SES.Score,High.Cholesterol,mean) #43.2 51.9

```

## Interpretation of other predictors
```{r}
#Interpretation of other predictors
#RaceWHITE:20.926
boxplot(Subject.SES.Score~Race) #low income leads to low Edu quality
boxplot(Edu~Race)
tapply(Edu,Race,mean) #14.4,15.7,2yr college vs 4yr college

```

## Should we use this one?
```{r}
m07<-lm(WASI~Race+Edu+SES+BMI*High.Cholesterol)
summary(m07) #R_adj=0.277, SES not significant and make cholesterl insignif
AIC(m07) #4317 vs AIC(m1)=4328
BIC(m07) #4355 vs BIC(m1)=4357
m08<-lm(WASI~Race+Edu+Subject.SES.Score+BMI*High.Cholesterol)
summary(m08) #R_adj=0.271, SES not significant and make cholesterl insignif
AIC(m08) #4330 vs AIC(m1)=4328
BIC(m08) #4363 vs BIC(m1)=4357
```

## Cross Validation
```{r}
library(dplyr)
training<-stroke %>% sample_n(.,length(stroke$Anonymized.ID)/2)
testing<-stroke[!(stroke$Anonymized.ID %in% training$Anonymized.ID),]
m6<-lm(WASI.Full.Scale...4.Percentile.Rank~Race+Volunteer.Highest.grade.completed.+BMI*High.Cholesterol,data=training)
summary(m6)
yhat <- m6 %>% predict(testing)
cor(yhat,testing$WASI.Full.Scale...4.Percentile.Rank) #0.475
plot(yhat,testing$WASI.Full.Scale...4.Percentile.Rank)
detach(stroke)
```


## Change to binary outcome
```{r}
median_WASI <- median(WASI)
WASI[WASI < median_WASI] <- 0
WASI[WASI >= median_WASI] <- 1
```


## Logistic regression
```{r}
m_logistic1 <-glm(WASI~Race+Edu+BMI+High.Cholesterol)
summary(m_logistic1)

round(exp(cbind(Estimate=coef(m_logistic1),confint(m_logistic1))),4)

```

## Interpretations:

Keeping all else constant, the odds of having an above median WASI score is 34.67% higher for those who are white

Keeping all else constant, for each year of increase in education, the odds of having an above median WASI score is increased by 5.69%

Keeping all else constant, for each point of increase in BMI, the odds of having an above median WASI score is decreased by 1.35%

Keeping all else constant, the odds of having an above median WASI score is 10.9% higher for those who have high cholesterol